---
title: "R Notebook"
output: github_document
---

```{r,eval=FALSE}
download.file("https://github.com/ANF-MetaBioDiv/course-material/archive/refs/heads/main.zip", destfile = "main.zip")
```

```{r}
refdb_folder <- here::here("data", "refdb")
refdb_folder
```

```{r}
if (!dir.exists(refdb_folder)) dir.create(refdb_folder, recursive = TRUE)
```

```{r}
getOption("timeout")
```
```{r}
# so we change timeout to be 20 minutes
options(timeout = 1200)

# we save in variable the path to the refdb
# in the working space
silva_train_set <- file.path(refdb_folder,
                             "silva_nr99_v138.1_train_set.fa.gz")

silva_species_assignment <- file.path(refdb_folder,
                                      "silva_species_assignment_v138.1.fa.gz")

# then we download the files if they don't already exist

if (!file.exists(silva_train_set)) {
  download.file(
    "https://zenodo.org/record/4587955/files/silva_nr99_v138.1_train_set.fa.gz",
    silva_train_set,
    quiet = TRUE
  )
}

if (!file.exists(silva_species_assignment)) {
  download.file(
    "https://zenodo.org/record/4587955/files/silva_species_assignment_v138.1.fa.gz",
    silva_species_assignment,
    quiet = TRUE
  )
}
```

```{r}
devtools::load_all(path="/Users/yvanohemartinier/ADM_tutoriel_dada2/course-material-main/R")
```

```{r}
path_to_fastqs <- here::here("data", "raw")
```

```{r}
fnFs <- sort(list.files(path_to_fastqs,
                        pattern = "_R1.fastq.gz",
                        full.names = TRUE))
```
#avec cette commande va chercher tous les noms de fichier "_R1.fastaq.gz" Mais on ne les veut pas tous

```{r}
fnRs <- sort(list.files(path_to_fastqs,
                        pattern = "_R2.fastq.gz",
                        full.names = TRUE))
```

## 2.2 Extract sample names
```{r}
sample_names <- basename(fnFs) |>
  strsplit(split = "_") |>
  sapply(head, 1)
```

#utilisation d'un plit, sample_names <- basename(fnFs) |> 
Dans basename on va rajouter une conditions, la suivante
strsplit(split = "_") peremt de decouper les noms de base qui sont séparé avec le "_" 
```{r}
basename(fnFs) |>
  head()
```
# avant découpage
```{r}
basename(fnFs) |>
  strsplit(split = "_") |>
  head()
```
# après découpage
```{r}
basename(fnFs) |>
  strsplit(split = "_") |>
  sapply(head, 1) |>
  head()
```

```{r}
gsub("^.+/|_.+$", "", fnFs) |> head()
```

## 3.Sequence quality check
```{r}
# create a directory for the outputs
quality_folder <- here::here("outputs",
                             "dada2",
                             "quality_plots")

if (!dir.exists(quality_folder)) {
  dir.create(quality_folder, recursive = TRUE)
}

qualityprofile(fnFs,
               fnRs,
               file.path(quality_folder, "quality_plots.pdf"))
```

## 4 Primer removla
4.1 Prepare outputs
```{r}
path_to_trimmed_reads <- here::here(
  "outputs",
  "dada2",
  "trimmed"
)

if (!dir.exists(path_to_trimmed_reads)) dir.create(path_to_trimmed_reads, recursive = TRUE)
```

4.2 Remove primers
```{r}
primer_fwd  <- "CCTACGGGNBGCASCAG"
primer_rev  <- "GACTACNVGGGTATCTAAT"
```

```{r}
Biostrings::readDNAStringSet(
  fnFs[1],
  format = "fastq",
  nrec = 10
)
```
```{r}
Biostrings::readDNAStringSet(
  fnRs[1],
  format = "fastq",
  nrec = 10
)

```



4.3 For more complex situations 